## Antibody Modelling Function for mice
## Create Fig 4-A


######### required packages ############
```{r}
required_packages <- c('dplyr', 'deSolve', 'ggplot2','purrr')

install_and_load <- function(packages) {
  for (pkg in packages) {
    if (!require(pkg, character.only = TRUE)) {  
      install.packages(pkg, dependencies = TRUE) 
      library(pkg, character.only = TRUE)        
    } else {
      library(pkg, character.only = TRUE)        
    }
  }
}

install_and_load(required_packages)
```

########### read parameter ############
```{r}
pop <- read.csv("populationParameters.txt",  row.names = 1)
```

######## Set time range ###########
```{r}
Tmin <- 0
Tmax <- 180 
step_size <- 0.005
times<-c(seq(Tmin,Tmax,step_size))
```

######### Simulate population trajectory ############
```{r}
## ODE Function to model population antibody dynamics
ode_ab_pop <- function(pars) {
  
  k1 <- as.numeric(pars[1]) # Growth rate in phase 1
  k2 <- as.numeric(pars[2]) # Decay rate in phase 2
  A0 <- as.numeric(pars[3]) # Initial condition
  C <- as.numeric(pars[4])  # Carrying capacity for sigmoid growth phase
  
  times <- c(seq(Tmin, Tmax, step_size))
  
  k_time <- data.frame(times = times, k = rep(0, length(times)))
  k_time <- k_time %>% mutate(k = case_when(times <= 35 ~ k1,  # set 35 days as the timing of peak
                                            times > 35 ~ -k2))
  
  k_t <- approxfun(k_time$times, k_time$k, rule = 2)
  
  derivs <- function(times, y, pars, k_t) {
    with(as.list(c(pars, y)), {
      k <- k_t(times)   # Get k value based on time
      if (times < 35) {
        dA <- k * A * (1 - A / C) # Sigmoid growth phase
      } else {
        dA <- k * A # Exponential decay phase
      }
      return(list(c(dA)))
    })
  }
  
  y <- c(A = A0) # Initial observation
  
  out <- ode(y = y, parms = pars, times = times, func = derivs, k_t = k_t)
  as.data.frame(out)
}
```

```{r}
## Sampling function to simulate the CI for population dynamics
sample_ab <- function(pop, num, vac){
  
  inds_mean <- which(rownames(pop) %in% c("k1_pop","k2_pop","A0_pop", 'peak_pop')) #row number of parameters
  inds_sd <- which(rownames(pop) %in% c("omega_k1","omega_k2","omega_A0", "omega_peak"))

  pars <- matrix(0, num+1, length(inds_mean))
  
  for (i in 1:length(inds_mean)) {
    mean_par <- pop$value[inds_mean[i]]
    sd_par <- pop$value[inds_sd[i]]
    #sd_par <- pop$se_sa[inds_mean[i]] ####standard error of population parameter
    
    if (i==1) {
      beta_k1_0_1ug <- pop$value[inds_mean[i]+1]
      beta_k1_0_2ug <- pop$value[inds_mean[i]+2]
      beta_k1_0_5ug <- pop$value[inds_mean[i]+3]
      beta_k1_1ug   <- pop$value[inds_mean[i]+4]
      log_k1 = log(mean_par)+(vac=='0.1ug Pfizer')*beta_k1_0_1ug+(vac=='0.2ug Pfizer')*beta_k1_0_2ug+
        (vac=='0.5ug Pfizer')*beta_k1_0_5ug + (vac=='1ug Pfizer')*beta_k1_1ug
      pars[1:num, i] <- exp(rnorm(num, mean = log_k1, sd=sd_par))
      pars[num+1, i] <- exp(log_k1)
    } 
    
    else if (i==2 ) {
      beta_k2_0_1ug <- pop$value[inds_mean[i]+1]
      beta_k2_0_2ug <- pop$value[inds_mean[i]+2]
      beta_k2_0_5ug <- pop$value[inds_mean[i]+3]
      beta_k2_1ug   <- pop$value[inds_mean[i]+4]
      log_k2 = log(mean_par)+(vac=='0.1ug Pfizer')*beta_k2_0_1ug+(vac=='0.2ug Pfizer')*beta_k2_0_2ug+
        (vac=='0.5ug Pfizer')*beta_k2_0_5ug +(vac=='1ug Pfizer')*beta_k2_1ug
      pars[1:num, i] <- exp(rnorm(num, mean = log_k2, sd=sd_par))
      pars[num+1, i] <- exp(log_k2)
    } 
    
    else if (i==3) {
      log_A0 = log(mean_par)
      pars[1:num, i] <- exp(rnorm(num, mean=log_A0, sd=sd_par))
      pars[num+1, i] <- exp(log_A0)
    }
    else if (i==4 ) {
      log_C = log(mean_par)
      pars[1:num, i] <- exp(rnorm(num, mean=log_C, sd=sd_par))
      pars[num+1, i] <- exp(log_C)
    } 
  }
  return(pars)
}
```

############# Function to store output in column form ##############
```{r}
run_ODE_control <- function(pars){
  total_AL <- matrix(NA,nrow=length(seq(Tmin,Tmax,step_size)),ncol=n)
  for(i in 1:n){
    out <- ode_ab_pop(pars[i, ])
  total_AL[,i] <- out$A
  }
  return(total_AL)
}
```

############## Introduce the covariate of DOSE for different mice groups ################
```{r}
df_cov <- data.frame(Vaccine = c( "0.05ug Pfizer", 
                                  "0.1ug Pfizer", 
                                  "0.2ug Pfizer", 
                                  "0.5ug Pfizer",
                                  "1ug Pfizer")
                 )

df_cov
```
############## Simulate the trajectory ####################
```{r}
n = 1000  # Can change to 100, or change stepsize to a larger value to save time

Fit <- list()
pb <- txtProgressBar(min = 0, max = 2, style = 3)
for (g in 1:5){
  vac <- df_cov$Vaccine[g]
  par <- c(k1=exp(log(pop["k1_pop", "value"])+(vac=='0.1ug Pfizer')*beta_k1_0_1ug+(vac=='0.2ug Pfizer')*beta_k1_0_2ug+(vac=='0.5ug Pfizer')*beta_k1_0_5ug+(vac=='1ug Pfizer')*beta_k1_1ug),
           k2=exp(log(pop["k2_pop", "value"])+(vac=='0.1ug Pfizer')*beta_k2_0_1ug+(vac=='0.2ug Pfizer')*beta_k2_0_2ug+(vac=='0.5ug Pfizer')*beta_k2_0_5ug+(vac=='1ug Pfizer')*beta_k2_1ug),
           A0=pop["A0_pop", "value"],
           C = pop["peak_pop", "value"])
  
  best_fit <- ode_ab_pop(par)
  
  pars <- sample_ab(pop, n, vac)
  total_VL <- run_ODE_control(pars)
  
  Min90  <- apply(total_VL,1,function(x){quantile(x,0.025,na.rm=T)})
  Max90  <- apply(total_VL,1,function(x){quantile(x,0.975,na.rm=T)})
  Min50  <- apply(total_VL,1,function(x){quantile(x,0.25,na.rm=T)})
  Max50  <- apply(total_VL,1,function(x){quantile(x,0.75,na.rm=T)})
  Mean50  <- apply(total_VL,1,function(x){quantile(x,0.5,na.rm=T)})
  index <- which(Mean50 < -2)[1]
  # Check if an index was found, and then replace the value
  if (!is.na(index)) {
    Mean50[index] <- -2
  }  
  
  Fit[[g]] <- cbind(best_fit,Min90,Max90,Min50,Max50,Mean50,vac)
  Fit[[g]] <- data.frame(Fit[[g]])
  colnames(Fit[[g]]) <- c("time","best_fit","Min95","Max95","Min50","Max50","Mean50","Group")
  setTxtProgressBar(pb, g)
}
close(pb)
combine_pop <- map_df(Fit, ~as.data.frame(.x))

```
############################
##### Figure 4-A ###########
############################

```{r}
# Create the plot
plot_pop1 <- ggplot(combine_pop, aes(x = time, y = best_fit)) +
  geom_ribbon(aes(ymin = Min95, ymax = Max95, group = Group, fill = Group), alpha = 0.1) +
  geom_ribbon(aes(ymin = Min50, ymax = Max50,group = Group, fill = Group), alpha = 0.15) +
  geom_line(aes(group = Group, color = Group),linewidth = 1.5) +
  labs(
    #title = "Visualization of Antibody Response",
   x = "days since first vaccination",
       y = "anti-spike IgG (%binding)",
    color = "Group",
    fill = "Group"
  ) +
   theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='bottom',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  )+ theme(
 # plot.title = element_text(size = 16),        
  axis.title.x = element_text(size = 14),       
  axis.title.y = element_text(size = 14),       
  axis.text = element_text(size = 14) ,         
  legend.title = element_text(size = 14),  
  legend.text = element_text(size = 12),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
)
plot_pop1
```

#########################################
######Individual Trajectory #############
#########################################

## read original data
```{r}
original <- read.csv("demo.csv") 
colnames(original)[colnames(original) == "Mice.Number"] <- "Code"
original
```
## load Monolix results ##
```{r}
Est <- read.csv("IndividualParameters/estimatedIndividualParameters.txt", sep = ",", comment.char = "", header = T)
Simulated <- read.csv("IndividualParameters/simulatedIndividualParameters.txt", sep = ",", comment.char = "", header = T)
max(Simulated$rep) #See how many simulated parameters for each individual
```
## Function 
```{r}
Tmin <- 0
Tmax <- 180 
step_size <- 0.01
times<-c(seq(Tmin,Tmax,step_size))
n <- 100

ind_fit_plt<-function(Est,Simulated, original){
  
  Fit <- list()
  for(i in 1:nrow(Est)){
    pars <- c(k1=Est$k1_mode[i],
              k2=Est$k2_mode[i],
              A0=Est$A0_mode[i],
              C = Est$peak_mode[i])
    fitted <- ode_ab_pop(pars)
    d1 <- data.frame(Days=(times),y=(fitted$A))
    Code <- Est$id[i]
    
    S <- 100 ###100 repeats
    P <- matrix(NA,nrow=length(times),ncol=S)
    
    
    for(j in 1:S){
      pars <- c(k1=Simulated$k1[j+S*(i-1)],k2=Simulated$k2[j+S*(i-1)],A0=Simulated$A0[j+S*(i-1)],C=Simulated$peak[j+S*(i-1)])
      out  <- ode_ab_pop(pars)
      P[,j] <- out$A
    }
    
    Min  <- apply(P,1,function(x){quantile(x,0.005)})
    Max  <- apply(P,1,function(x){quantile(x,0.995)})
    
    fit <- cbind(d1,Min,Max,Code)
    fit$Max[fit$Max> 100] <- 100
    fit$Min[fit$Min> 100] <- 100
    
    Fit[[i]] <- data.frame(fit)
  }
  
  ind_fit_unlist <- map_df(Fit, ~as.data.frame(.x))
  unique_Vacc <- original[!duplicated(original$Code), c("Code", "Vaccine.Group")]
  ind_fit <- merge(ind_fit_unlist, unique_Vacc, by="Code",all.x=TRUE) 
  ind_fit$Code <- as.factor(ind_fit$Code)
  ind_fit <- ind_fit[order(ind_fit$Code, ind_fit$Days), ]
  return(ind_fit)
}

individual_fit <- ind_fit_plt(Est, Simulated, original)
```
######################################
#### Supplementary Figure 3 ##########
######################################

```{r}
plot_indfit <- ggplot(individual_fit, aes(x = Days)) +
  # Add shaded area for Min/Max range
  geom_ribbon(aes(ymin = Min, ymax = Max, fill = Vaccine.Group), alpha = 0.3) +
  # Add the fit line
  geom_line(aes(y = y, color = Vaccine.Group), size = 1, alpha = 0.8) +
  # Add individual's point
  geom_point(data = original, aes(x = Days, y = Antibody.Level, color = Vaccine.Group), 
             shape = 16, size = 2, alpha = 2) +
  # Separate plots for each individual
  facet_wrap(~ as.factor(Code)) +
  # Add labels and theme
  labs(
    #title = "Individual Fit with Shaded Uncertainty",
     x = "days since first vaccination",
       y = "anti-spike IgG (%binding)",
    fill = "Vaccine Group",
    color = "Vaccine Group"
  ) +
 theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='bottom',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  ) + theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 12),        # X-axis label font size
  axis.title.y = element_text(size = 12),        # Y-axis label font size
  axis.text = element_text(size = 12) ,           # Axis tick labels font size
  legend.title = element_text(size = 12),  # Legend title font size
  legend.text = element_text(size = 10),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
)

plot_indfit
```




